Referring Expressions as Formulas of Description Logic

### 2009-01-21 trunk/Papers/dl-gre-jair/discussion.tex
jair article: only reformatted inlg article

\begin{figure}
\begin{center}
\includegraphics[width=0.2\textwidth]{drawers.png}
\end{center}\vspace*{-3ex}
\caption{A schematic view of the filing cabinets.}\label{fig:drawers}\vspace*{-1.5ex}
\end{figure}

\newcite{viethen06:_algor_for_gener_refer_expres} describe the data in
more detail and present results of evaluating the Full Brevity
algorithm, the Incremental Algorithm (both by \newcite{Dale1995}), and
the Relational Algorithm \cite{dale91:_gener_refer_expres_invol_relat}
on this corpus. The Incremental Algorithm is dependent on a predefined
ordering in which properties are added to the description. Viethen and
Dale, therefore, try all possible orderings and evaluate what
percentage of descriptions an algorithm can generate with any of
them. The Full Brevity and the Relational Algorithms choose properties
based on their discriminatory power and only use the orderings as tie
breakers. Viethen and Dale found that the Incremental Algorithm is
capable of generating 98 of the 103 non-relational
descriptions. However, the Relational Algorithm was unable to
generate even a single one of the human-generated relational
descriptions.

@@

When we add relations, our algorithm is able to generate 10 of the 15
human-produced relational descriptions correctly (in addition to the
98 non-relational descriptions).  Fig.~\ref{fig:example_outputs} gives
example outputs of the \el\ algorithm for three different orderings,
which together achieve this coverage.  Of the five human-produced
descriptions that the \el\ algorithm cannot generate, three involve
references to sets (\textit{the two blues ones in horizontal
  sequence}/\textit{the two yellow drawers}), and two contain so much
redundant information that our algorithm cannot reproduce them:
Similarly to the Incremental Algorithm, it allows for some redundancy,
but stops once it has found a distinguishing description. It does,
however, generate other, simpler descriptions for these three
referents. 
%human-produced descriptions that cannot be reproduced exactly.

%two mention column information, which we excluded from our domain, and

@@

Our GRE algorithms do not guarantee that the formula they compute can
actually be realized in language.  For example, none of the formulas
our algorithms computed in the Viethen and Dale domain contained an
atom that would commonly be realized as a noun; the property
$\mathsf{drawer}$ is never used because it applies to all individuals
in the domain.  This particular problem could easily be worked around
in a post-processing step.  However, another problem arises from the
massive use of negation in the \alc\ algorithm; it will be hard for
any realizer to find a reasonable way of expressing a formula like
$\neg \exists R.(\neg P \sqcap \neg Q)$ as a smooth noun phrase.
Although we agree with \newcite{deemter02:_gener_refer_expres} and
others that the careful use of negation and disjunction can improve
REs, these connectives must not be overused.  Thus we consider the
formulas computed by the \el\ algorithm ``safer'' with respect to
realization.

Of course, we share the problem of interfacing GRE and realization
with every other approach that separates these two modules, i.e.,
almost the entire GRE literature (notable exceptions are, e.g.,~\newcite{Horacek1997} and SPUD \cite{Stone1998a}).  In principle, we
believe that it is a good idea to handle sentence planning and
realization in a single module; for instance, SPUD can use its
awareness of the syntactic context to generate succinct REs as in
``take the rabbit from the hat''.  We hope that the ideas we have
explored here for efficient and expressive RE generation can
eventually be combined with recent efficient algorithms for integrated
sentence planning and realization, such as in \newcite{KolSto07}.

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{runtimes}\vspace*{-2ex}
  \caption{Average runtimes (in ms) of the two algorithms on random
    models with different numbers of individuals.}
  \label{fig:runtimes}\vspace*{-1.5ex}
\end{figure}

>>>>

\begin{figure}
\begin{center}
\todo{need to reformat pic}
%%%\includegraphics[width=0.2\textwidth]{drawers.png}
\end{center}\vspace*{-3ex}
\caption{A schematic view of the filing cabinets.}\label{fig:drawers}\vspace*{-1.5ex}
\end{figure}

\citeA{viethen06:_algor_for_gener_refer_expres} describe the data in
more detail and present results of evaluating the Full Brevity
algorithm, the Incremental Algorithm (both by \citeA{Dale1995}), and
the Relational Algorithm \cite{dale91:_gener_refer_expres_invol_relat}
on this corpus. The Incremental Algorithm is dependent on a predefined
ordering in which properties are added to the description. Viethen and
Dale, therefore, try all possible orderings and evaluate what
percentage of descriptions an algorithm can generate with any of
them. The Full Brevity and the Relational Algorithms choose properties
based on their discriminatory power and only use the orderings as tie
breakers. Viethen and Dale found that the Incremental Algorithm is
capable of generating 98 of the 103 non-relational
descriptions. However, the Relational Algorithm was unable to
generate even a single one of the human-generated relational
descriptions.

@@

When we add relations, our algorithm is able to generate 10 of the 15
human-produced relational descriptions correctly (in addition to the
98 non-relational descriptions).  Fig.~\ref{fig:example_outputs} gives
example outputs of the \el\ algorithm for three different orderings,
which together achieve this coverage.  Of the five human-produced
descriptions that the \el\ algorithm cannot generate, three involve
references to sets (\textit{the two blues ones in horizontal
sequence}/\textit{the two yellow drawers}), and two contain so much
redundant information that our algorithm cannot reproduce them:
Similarly to the Incremental Algorithm, our algorithm allows for some
redundancy, but stops once it has found a distinguishing
description. It does, however, generate other, simpler descriptions
for these referents.

%these three referents.
%human-produced descriptions that cannot be reproduced exactly.
%two mention column information, which we excluded from our domain, and

@@

Our GRE algorithms do not guarantee that the formula they compute can
actually be realized in language.  For example, none of the formulas
our algorithms computed in the Viethen and Dale domain contained an
atom that would commonly be realized as a noun; the property
$\mathsf{drawer}$ is never used because it applies to all individuals
in the domain.  This particular problem could easily be worked around
in a post-processing step.  However, another problem arises from the
massive use of negation in the \alc\ algorithm; it will be hard for
any realizer to find a reasonable way of expressing a formula like
$\neg \exists R.(\neg P \sqcap \neg Q)$ as a smooth noun phrase.
Although we agree with \citeA{deemter02:_gener_refer_expres} and
others that the careful use of negation and disjunction can improve
REs, these connectives must not be overused.  Thus we consider the
formulas computed by the \el\ algorithm ``safer'' with respect to
realization.

Of course, we share the problem of interfacing GRE and realization
with every other approach that separates these two modules, i.e.,
almost the entire GRE literature (notable exceptions are, e.g.,~\citeA{Horacek1997} and SPUD \cite{Stone1998a}).  In principle, we
believe that it is a good idea to handle sentence planning and
realization in a single module; for instance, SPUD can use its
awareness of the syntactic context to generate succinct REs as in
``take the rabbit from the hat''.  We hope that the ideas we have
explored here for efficient and expressive RE generation can
eventually be combined with recent efficient algorithms for integrated
sentence planning and realization, such as in \citeA{KolSto07}.

\begin{figure}[t]
  \centering
	\todo{need to reformat pic}
  %%%\includegraphics[width=\columnwidth]{runtimes}\vspace*{-2ex}
  \caption{Average runtimes (in ms) of the two algorithms on random
    models with different numbers of individuals.}
  \label{fig:runtimes}\vspace*{-1.5ex}
\end{figure}

### 2009-01-21 trunk/Papers/dl-gre-jair/gre.tex
jair article: only reformatted inlg article

In the examples above, it is because $\mathsf{flower} \sqcap \exists
\mathsf{in}. \mathsf{hat}$ denotes exactly $\{f_2\}$ that we can say
``the flower in the hat'' to refer to $f_2$.  This perspective
provides a general framework into which many existing GRE approaches
fit: Traditional attribute selection \cite{Dale1995} corresponds to
building DL formulas that are conjunctions of atoms; relational REs as
in \newcite{dale91:_gener_refer_expres_invol_relat} are formulas of
\el; and so on.  We will further pursue the idea of organizing GRE
approaches with respect to the variant of DL they use in
Section~\ref{sec:related}.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{pic-dale-haddock}
  \caption{(a) The \newcite{dale91:_gener_refer_expres_invol_relat}
    scenario; (b) the \newcite{Stone1998a} scenario.}
  \label{fig:dale-haddock}\vspace*{-1ex}
\end{figure}

@@

It can be shown that for \alc, the similarity sets of a finite model
coincide exactly with the \emph{simulation classes} of this model.
Simulation classes have been studied extensively in the literature
(see e.g.,~\newcite{blac:moda01}; \newcite{kurt:expr99}), and there are several
efficient algorithms for computing \alc-simulation classes
\cite{hopc:algo71,paig:thre87,dovier04:_effic_algor_for_comput_bisim_equiv}.
However, these algorithms will only compute the simulation classes
themselves. Here we extend the \newcite{hopc:algo71} algorithm such
that it computes, along with each set, also a formula that denotes
exactly this set.  We can then use these formulas as representations
of the referring expressions.

@@

Let's try our algorithms on some examples.  We first run the \el\
algorithm on the model shown in Fig.~\ref{fig:dale-haddock}a, which is
taken from \newcite{dale91:_gener_refer_expres_invol_relat}.  The
algorithm starts with $\RE = \{\top\}$.  In the first loop, it adds
the formulas $\mathsf{floor}$, $\mathsf{bowl}$, $\mathsf{cup}$, and
$\mathsf{table}$, and then removes $\top$ because it is now subsumed.
Not all of these formulas denote singletons; for instance,
$\interp{\mathsf{cup}}$ contains two individuals.  So we iterate over
the relations to refine our formulas.  After the first iteration over
the relations, we have $\RE = \{ \mathsf{floor}, \mathsf{bowl} \sqcap
\exists \mathsf{on}.\mathsf{floor}, \mathsf{bowl} \sqcap \exists
\mathsf{on}.\mathsf{table}, \mathsf{cup}, \mathsf{table} \}$. Notice
that $\mathsf{bowl}$ has become subsumed, but we haven't distinguished
the cups and tables further.

>>>>

In the examples above, it is because $\mathsf{flower} \sqcap \exists
\mathsf{in}. \mathsf{hat}$ denotes exactly $\{f_2\}$ that we can say
``the flower in the hat'' to refer to $f_2$.  This perspective
provides a general framework into which many existing GRE approaches
fit: Traditional attribute selection \cite{Dale1995} corresponds to
building DL formulas that are conjunctions of atoms; relational REs as
in \citeA{dale91:_gener_refer_expres_invol_relat} are formulas of
\el; and so on.  We will further pursue the idea of organizing GRE
approaches with respect to the variant of DL they use in
Section~\ref{sec:related}.

@@

\begin{figure}[t]
  \centering
	\todo{need to reformat pic}
  %%%%%%%%%%\includegraphics[width=\columnwidth]{pic-dale-haddock}
  \caption{(a) The \citeA{dale91:_gener_refer_expres_invol_relat}
    scenario; (b) the \citeA{Stone1998a} scenario.}
  \label{fig:dale-haddock}\vspace*{-1ex}
\end{figure}

@@

It can be shown that for \alc, the similarity sets of a finite model
coincide exactly with the \emph{simulation classes} of this model.
Simulation classes have been studied extensively in the literature
(see e.g.,~\citeA{blac:moda01}; \citeA{kurt:expr99}), and there are several
efficient algorithms for computing \alc-simulation classes
\cite{hopc:algo71,paig:thre87,dovier04:_effic_algor_for_comput_bisim_equiv}.
However, these algorithms will only compute the simulation classes
themselves. Here we extend the \citeA{hopc:algo71} algorithm such
that it computes, along with each set, also a formula that denotes
exactly this set.  We can then use these formulas as representations
of the referring expressions.

@@

Let's try our algorithms on some examples.  We first run the \el\
algorithm on the model shown in Fig.~\ref{fig:dale-haddock}a, which is
taken from \citeA{dale91:_gener_refer_expres_invol_relat}.  The
algorithm starts with $\RE = \{\top\}$.  In the first loop, it adds
the formulas $\mathsf{floor}$, $\mathsf{bowl}$, $\mathsf{cup}$, and
$\mathsf{table}$, and then removes $\top$ because it is now subsumed.
Not all of these formulas denote singletons; for instance,
$\interp{\mathsf{cup}}$ contains two individuals.  So we iterate over
the relations to refine our formulas.  After the first iteration over
the relations, we have $\RE = \{ \mathsf{floor}, \mathsf{bowl} \sqcap
\exists \mathsf{on}.\mathsf{floor}, \mathsf{bowl} \sqcap \exists
\mathsf{on}.\mathsf{table}, \mathsf{cup}, \mathsf{table} \}$. Notice
that $\mathsf{bowl}$ has become subsumed, but we haven't distinguished
the cups and tables further.
